---
title: "NISTeHandbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NISTeHandbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



# Introduction to package {#intro}

The NISTeHandbook R package seeks to implement methods from the NIST/SEMATECH e-Handbook of Statistical Methods. 

As of the current version only 4 methods have been implemented:

* [4-plot](https://www.itl.nist.gov/div898/handbook/eda/section2/eda23.htm)

* [6-plot](https://www.itl.nist.gov/div898/handbook/eda/section3/6plot.htm)

* [Tietjen-Moore Test for outliers](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h2.htm)

* [Probability Plot Correlation Coefficient Plot](https://www.itl.nist.gov/div898/handbook/eda/section3/ppccplot.htm)


# 4-plot

4-plot and 6-plot are very similar EDA methods for testing assumptions on a given data.


Let us start with the 4 plot, using the example from the eNist website. We use the [LEW](https://www.itl.nist.gov/div898/handbook/datasets/LEW.DAT) dataset, which [describes](https://www.itl.nist.gov/div898/strd/univ/addinfo/lew.html) a study of the physical behaviour of steel-concrete beams under periodic load. The response variable is deflection from a center point, and 200 observations were collected.

We first load the RPackage:
```{r}
library(NISTeHandbook)
```

The package provides the data set directly.

```{r}
head(LEW.DAT)
```

To use the 4-plot function we simply provide the observations to the function:

```{r}
fourplots <- fourPlot(LEW.DAT$Deflection)
fourplots
```


The function outputs a 4plot object that contains every plot as a ggplot object, such that you can modify each individual plot. Printing the 4plot object gives the plot shown above. The number of bins is set to a default of 11, but we can specify other bin sizes in the 4plot function.

```{r}
fourplots <- fourPlot(LEW.DAT$Deflection, bins = 20)
```

## Using individual plots:

To use only a specific or modify plots we can access them in the 4plot object. Let us take the sequence plot.

```{r}
fourplots$seq_plot
```

If we wish to modify any of the plots we can use ggplot2:

```{r}
fourplots$seq_plot + geom_line(color="red")
```
For finding out the specific geoms of fourplot, see the documentation in ?fourplots.

We can then modify and save these to change our 4plot.

```{r}
fourplots$seq_plot <- fourplots$seq_plot + geom_line(color="red")
fourplots
```

For assumptions and conclusions of the plots you should consult the handbook.

# 6-plot

The 6-plot behaves in a very similar way to the 4-plot. The function attempts to do a linear Y versus X fit. Let us again load data from the handbook website, in another fashion:

```{r}
PONTIUS <- read.table("https://www.itl.nist.gov/div898/handbook/datasets/PONTIUS.DAT", skip=25)
head(PONTIUS)
```

We are looking at a response variable, beam deflections, V1 at various different loads V2. To create the six plot we input the response and the explanatory variable.

```{r}
sixPlot(PONTIUS$V2, PONTIUS$V1)
```


sixPlot returns a sixPlot object that contains the 6 ggplots, that can be manipulated in a similar manner to the 4plot. 



# Tietjen-Moore Test for outliers

Imagine having approximately normal distributed data, and you want to check for outliers. Then the Tietjen-Moore test is suited for you. You can use the TMTest to test for exactly $k$ outliers, no more, no less. 

Let us again take data from the [handbook](https://www.itl.nist.gov/div898/handbook/datasets/TIETMOO1.DAT) (Which Tietjen and Moore also used in the original paper): 

```{r}
x = c(-1.40, -0.44, -0.30, -0.24, -0.22, -0.13, -0.05, 0.06, 0.10, 0.18, 0.20, 0.39, 0.48, 0.63, 1.01)
```

The x-values are observations of vertical semi-diameters of the planet Venus...

After doing some EDA, we suspect there might be two outliers.


```{r}
TMTest(x,2)
```

The TMTest function shows a normal QQ plot, to remind us that the test is only valid on approximately normal distributed data. We also see the two suspected outliers.

The print shows us that we can conclude that the suspected outliers are indeed outliers.


# Tukey-Lambda PPCC Plot

